---
title: "Project - France"
author: "Jakub Doležal"
date: "15. 3. 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

#### Outline:

1.  Transformation to logs\
2.  Seasonal adjustments\
3.  Transformation to percentage changes\
4.  Estimation, diagnostics and forecasting\
5.  Output gap

#### Preliminaries

Libraries

```{r}
rm(list=ls())

library(repr)
library(uroot)
library(urca)
library(strucchange)
library(stats)
library(pdfetch)
library(xts)
library(ggplot2)
library(ggfortify) 
library(forecast)
library(tsbox) 
library(seasonal)
library(tseries)
library(TSA)
library(mFilter)
library(neverhpfilter)
```

**Data:**

We use the following units:

GDP: Chain linked volumes (2015) HICP: 2015=100

```{r}
gdp_xts_q <- pdfetch_EUROSTAT("namq_10_gdp", S_ADJ="SCA", NA_ITEM="B1GQ", FREQ="Q", UNIT="CLV15_MEUR", GEO="FR")

inf_xts_m <- pdfetch_EUROSTAT("PRC_HICP_MIDX", CL_COICOP="CP00", FREQ="M", UNIT="I15", GEO="FR")

inf_xts_m00 <- inf_xts_m[,"M.I15.CP00.FR"]

```

Conversion to ts object:

```{r}
gdp <- ts_first_of_period(gdp_xts_q)
gdp_ts <- ts_ts(gdp)
gdp <- gdp_ts

inf <- ts_first_of_period(inf_xts_m00)
inf_ts <- ts_ts(inf)
inf <- inf_ts
```

### 1. Tranformation to logs

```{r}
log_gdp <- log(gdp)
log_inf <- log(inf)

autoplot(log_gdp, main = "log of real GDP (chain linked volumes (2015))")
autoplot(log_inf, main = "log of HICP (2015=100)")
```

### 2. Seasonal adjustments

```{r}

log_inf

linf <- window(log_inf, end=c(2023,3))

inf_ts_log_seas <- seas(linf)
inf_log_seas <- inf_ts_log_seas$data[,1]

autoplot(inf_ts_log_seas)
```

### 3. Transformation to percentage changes

Transformed series of GDP are indicated as stationary by ADF test. However, log-difference is not enough to make HICP in both cases stationary. We therefore take simple log-difference and address this issues in ARIMA model (p, 1, q) as suggested in the feedback\

From now on:\
gdp_ld = log-differenced q-o-q GDP\
inf_ld = log-differenced m-o-m HICP\
gdp_yoy = log-differenced y-o-y GDP\
inf_yoy = log-differenced y-o-y HICP

```{r}
gdp_ld <- diff(log_gdp)*100 #q-o-q
inf_ld <- diff(inf_log_seas)*100 #m-o-m

?arima

gdp_yoy <- na.omit(diff(log_gdp, lag = 4)*100)
inf_yoy <- na.omit(diff(inf_log_seas, lag = 12)*100)

adf.test(gdp_ld)
adf.test(inf_ld)
adf.test(gdp_yoy)
adf.test(inf_yoy)

autoplot(gdp_ld)
autoplot(inf_ld)
autoplot(gdp_yoy)
autoplot(inf_yoy)
```

Periodogram and Spectrum Analysis: Periodogram does not show any sign of seasonality. Our time series are I(1), except the m-o-m HICP and y-o-y HICP (I(2)).

```{r}
periodogram(gdp_ld) #periodogram
spec.pgram(gdp_ld, detrend = FALSE) #spectrum without detrend

periodogram(inf_ld) #periodogram
spec.pgram(inf_ld, detrend = FALSE) #spectrum without detrend
```

Structrural Breaks:

Z-A test identifies a structural break both in GDP (attributed to COVID-19 pandemic) and HICP (record-high inflation of the past months) in the very end of our data set. However, no structural breaks are found by the Chow test.

```{r}
chow_inf <- inf_ld
plot.ts(chow_inf)
sctest(ts(chow_inf[1:325]) ~ chow_inf[2:326], type="Chow")

chow_gdp <- gdp_ld
plot.ts(chow_gdp)
sctest(ts(chow_gdp[1:171]) ~ chow_gdp[2:172], type="Chow")
```

```{r}
za_GDP <- ur.za(gdp_ld, model = c("intercept", "trend", "both"), lag=NULL)
summary(za_GDP)
plot(za_GDP)

za_inf <- ur.za(inf_ld, model = c("intercept", "trend", "both"), lag=NULL)
summary(za_inf)
plot(za_inf)
```

Identification Correlogram: We perform the Box-Jenkins step-by-step estimation consisting of identification, estimation, diagnostics and forecasting.

```{r}
acf(gdp_ld, lag.max = 24)
pacf(gdp_ld, lag.max = 24)

acf(inf_ld, lag.max = 24)
pacf(inf_ld, lag.max = 24)

acf(gdp_yoy, lag.max = 24)
pacf(gdp_yoy, lag.max = 24)

acf(inf_yoy, lag.max = 24)
pacf(inf_yoy, lag.max = 24)

```

### 4. Estimation, diagnostics and forecasting

Estimation:

We construct ARIMA models according to ACFs and PACFs. For HICP, the model takes (p, 1, q) for to ensure stationarity of estimated series.

```{r}
fit_gdp <- forecast::Arima(gdp_ld, order = c(2,0,1), include.mean = TRUE)
fit_gdp

fit_inf <- forecast::Arima(inf_ld, order = c(1,1,4), include.mean = TRUE)
fit_inf

fit_gdp_yoy <- forecast::Arima(gdp_yoy, order = c(1,0,3), include.mean = FALSE)
fit_gdp_yoy

fit_inf_yoy <- forecast::Arima(inf_yoy, order = c(2,1,4), include.mean = FALSE)
fit_inf_yoy
```

Diagnostics:

We employed the sarima function from "astsa" package for complex diagnostics. Residuals are uncorrelated, normally distributed and thus resemble white noise, which is confirmed by the Ljung-Box test (we are unable to reject the iid nature of residuals).

```{r}
library(astsa)
fit_gdp_diag <- sarima(gdp_ld, p = 2, d = 0, q = 1, no.constant = FALSE, details = TRUE) #diagnostics ARIMA(2,0,1)

fit_inf_diag <- sarima(inf_ld, p = 1, d = 1, q = 4, no.constant = FALSE, details = TRUE) #diagnostics ARIMA(1,1,4)

fit_gdp_yoy_diag <- sarima(gdp_yoy, p = 1, d = 0, q = 3, no.constant = TRUE, details = TRUE) #diagnostics ARIMA(1,0,2)

fit_inf_yoy_diag <- sarima(inf_yoy, p = 2, d = 1, q = 4, no.constant = TRUE, details = TRUE) #diagnostics ARIMA(2,1,4)
```

Forecasting:

```{r}
fcst_gdp <- forecast::forecast(fit_gdp, h=8)
plot(fcst_gdp)

fcst_inf <- forecast::forecast(fit_inf, h=24)
plot(fcst_inf)

fcst_gdp_yoy <- forecast::forecast(fit_gdp_yoy, h=8)
plot(fcst_gdp_yoy)

fcst_inf_yoy <- forecast::forecast(fit_inf_yoy, h=24)
plot(fcst_inf_yoy)
```

### 5. Output gap

Hodrick-Prescott filter:

```{r}
hp <- hpfilter(log_gdp, freq = 1600)
hp_trend <- hp$trend
hp_cycle <- hp$cycle
plot(hp)
```

Hamilton filter:

# Hamilton filter

```{r}
# Requires xts format
log_gdp_xts <- ts_xts(log_gdp)
gdp_HAM <- yth_filter(log_gdp_xts, h= 8, p = 4)
par(mfrow=c(1,1))
plot(gdp_HAM["1980/2022"][,1:2], grid.col = "white", legend.loc = "topleft", main = "Log of Real GDP and trend", panels = 'lines(gdp_HAM["1980/2022"][,3], type="h", on=NA)')
HAM_trend <- ts_ts(gdp_HAM$value.trend)
HAM_cycle <- ts_ts(gdp_HAM$value.cycle)
```

Spectra comparison:

Output gaps are dominantly defined by low frequencies (trends) as shown by the spectrum analysis. On the other hand GDP is defined mostly by higher frequencies (probably a result of using log-differenced time series).

```{r}
library(TSA)

periodogram(gdp_ld) #periodogram
spec.pgram(gdp_ld, detrend = FALSE) #spectrum without detrend

periodogram(gdp_yoy) #periodogram
spec.pgram(gdp_yoy, detrend = FALSE) #spectrum without detrend

periodogram(hp_cycle) #periodogram
spec.pgram(hp_cycle, detrend = FALSE) #spectrum without detrend

HAM_per <- window(HAM_cycle, start = c(1982,4))

periodogram(HAM_per) #periodogram
spec.pgram(HAM_per, detrend = FALSE) #spectrum without detrend
```

Stationarity and forecast: Both output gap time series are stationary.

```{r}
adf.test(hp_cycle)
adf.test(HAM_per)

fit_hp <- auto.arima(hp_cycle, seasonal = FALSE)
fit_HAM <- auto.arima(HAM_per, seasonal = FALSE)

fcst_hp <- forecast::forecast(fit_hp, h=8)
fcst_HAM <- forecast::forecast(fit_HAM, h=8)

plot(fcst_hp)
plot(fcst_HAM)
```
